### 2021-06-03 JSS/using_drtmle_jss.Rmd
add jss files

---
documentclass: jss
author:
  - name: David Benkeser
    affiliation: Emory University
    address: |
      | 1518 Clifton Road, NE
      | Mailstop: 002-3AA
      | Atlanta, Georgia, 30322, U.S.A.
    email: \email{benkeser@emory.edu}
    orcdi: 0000-0002-1019-8343
  - name: Nima S. Hejazi
    affiliation: University of California, Berkeley
    email: \email{nhejazi@berkeley.edu}
    orcid: 0000-0002-7127-2789
title:
  formatted: "Doubly-Robust Inference in R using \\pkg{drtmle}"
  plain: "Doubly-Robust Inference in R using drtmle"
  short: "\\pkg{drtmle}: Doubly-robust Inference"
keywords:
  formatted: [causal inference, machine learning, treatment effects, targeted minimum loss estimation]
  plain: [causal inference, machine learning, treatment effects, targeted minimum loss estimation]
preamble: >
  \usepackage{amsmath}
abstract: >
  Inverse probability of treatment weighted estimators and doubly-robust
  estimators (including augmented inverse probability of treatment weight and
  targeted minimum loss-based estimators) are widely used in causal inference to
  estimate and draw inference about the average effect of a treatment. As an
  intermediate step, these estimators require estimation of key nuisance
  parameters, which are often regression functions. Typically, regressions are
  estimated using maximum likelihood and parametric models. Confidence intervals
  and p-values may be computed based on standard asymptotic results, such as the
  central limit theorem, the delta method, and the nonparametric bootstrap.
  However, in high-dimensional settings maximum likelihood estimation often
  breaks down and standard procedures no longer yield correct inference.
  Instead, we may rely on adaptive estimators of nuisance parameters to
  construct flexible regression estimators.  However, use of adaptive estimators
  poses a challenge for performing statistical inference about an estimated
  treatment effect. While doubly-robust estimators facilitate inference when
  *all* relevant regression functions are consistently estimated, the same
  cannot be said when at least one estimator is inconsistent. \pkg{drtmle}
  implements doubly-robust confidence intervals and hypothesis tests about
  targeted minimum loss-based estimates of the average treatment effect. The
  package additionally implements an inverse probability of treatment weighted
  estimator that yields valid inference even when the probability of treatment
  is estimated via adaptive methods.
output: rticles::jss_article
bibliography: refs
---

@@

In many fields researchers are interested in assessing the population-level
effect of a particular treatment on an outcome of interest. The treatment might
correspond to a drug, a potentially harmful exposure, or a policy intervention.
Often, the treatment cannot be randomized due to ethical or logistical reasons.
This has sparked great interest in developing statistical methodology to
estimate population-level effects from observational data.  Estimation of these
effects often requires accounting for putative confounding factors, that is,
factors related to whether a participant receives treatment and to the
participant's outcome. In many settings, researchers measure many potential
confounders, for example using administrative databases or genetic sequencing
technology. Due to the curse-of-dimensionality, common statistical estimation
techniques are not feasible for such high-dimensional data without restrictive
modeling assumptions. When these assumptions are violated, estimates of the
population-level effect of a treatment may have large bias.

>>>>

---
documentclass: jss
author:
  - name: David Benkeser
    affiliation: Emory University
    address: |
      | 1518 Clifton Road, NE
      | Mailstop: 002-3AA
      | Atlanta, Georgia, 30322, U.S.A.
    email: \email{benkeser@emory.edu}
  - name: Nima Hejazi
    affiliation: University of California, Berkeley
    email: \email{nhejazi@berkeley.edu}
title:
  formatted: "Doubly-Robust Inference in R using \\pkg{drtmle}"
  plain: "Doubly-Robust Inference in R using drtmle"
  short: "\\pkg{drtmle}: Doubly-robust Inference"
keywords:
  formatted: [causal inference, machine learning, treatment effects, targeted minimum loss estimation]
  plain: [causal inference, machine learning, treatment effects, targeted minimum loss estimation]
preamble: >
  \usepackage{amsmath}
abstract: >
  Inverse probability of treatment weighted estimators and doubly-robust
  estimators (including augmented inverse probability of treatment weight and
  targeted minimum loss-based estimators) are widely used in causal inference to
  estimate and draw inference about the average effect of a treatment. As an
  intermediate step, these estimators require estimation of key nuisance
  parameters, which are often regression functions. Typically, regressions are
  estimated using maximum likelihood and parametric models. Confidence intervals
  and p-values may be computed based on standard asymptotic results, such as the
  central limit theorem, the delta method, and the nonparametric bootstrap. 
  However, in high-dimensional settings maximum likelihood estimation often breaks 
  down and standard procedures no longer yield correct inference. Instead, we may 
  rely on adaptive estimators of nuisance parameters to construct flexible regression estimators.
  However, use of adaptive estimators poses a challenge for performing
  statistical inference about an estimated treatment effect. While doubly-robust
  estimators facilitate inference when *all* relevant regression functions are
  consistently estimated, the same cannot be said when at least one estimator is
  inconsistent. \pkg{drtmle} implements
  doubly-robust confidence intervals and hypothesis tests about targeted minimum
  loss-based estimates of the average treatment effect. The package additionally
  implements an inverse probability of treatment weighted estimator that yields
  valid inference even when the probability of treatment is estimated via
  adaptive methods.
output: rticles::jss_article
bibliography: refs
---

@@

In many fields researchers are interested in assessing the population-level
effect of a particular treatment on an outcome of interest. The treatment might
correspond to a drug, a potentially harmful exposure exposure, or a policy
intervention. Often, the treatment may not be randomized due to ethical or
logistical reasons. This has sparked great interest in developing statistical
methodology to estimate population-level effects from observational data.
Estimation of these effects often requires accounting for putative confounding
factors, that is, factors related to whether a participant receives treatment
and to the participant's outcome. In many settings, researchers measure many
potential confounders, for example using administrative databases or genetic
sequencing technology. Due to the curse-of-dimensionality, common statistical
estimation techniques are not feasible for such high-dimensional data without
restrictive modeling assumptions. When these assumptions are violated,
estimates of the population-level effect of a treatment may have large bias.
